\section{the language}
Creeve (Krivittè) is a fictional language spoken in an approximate geographical region Kritenno by the fictional anthropomorphic vulpes that settle mainly around the estuary of the end of Iallve river. Creeve is thought to be part of a wider group of languages spoken around the Kritenno known as Eivas language family (from \emph{eivè}, wind, air, or gusts). Because of the extensive area where Creeve is spoken, many pockets of dialect formed as a spectrum or transition which has been assimilated by another language. Hence, the variety of Creeve described here can be called standard Creeve. This can loosely be identified as proper Creeve and language of unification.

Compared to another group of language, Southeastern Eivas are thought to be closer to the ancestral language than others. It is assumed that the phenomenon is caused by their beings relatively isolated to others, with a few points and plains separated them with Allerrene speaker on the east. Meanwhile, Coastal Eivas languages are heavily influenced by the mainstream activity of the sailings and trade driven by the exports of azurite, salt, and crafts. The great exposure of the outside causes Coastal Eivas tends to contain more loanwords and not conservative and regarded as not proper by the western and southern language speaker.

Further, Creeve is in the state of rapid changes. New words and form of new colloquial forms could suddenly appear; some to acquire permanent acceptance, others vanish. That is because the condition of where Creeve is spoken. Iallve estuary is a major trade point of northern civilization and the mineral tycoon from Aveima and the east. With enormous influence on the language itself, structures regarded as improper could be mainstream and become acceptable in short time. Because of the rapid changes, this grammar has no intention to capture diachronic changes, but rather to present a synchronic statement.

Speakers of Creeve generally refer themselves as \emph{leirì vimmi,} `people of the salt plain'. The name of the language itself is unknown by the speaker. Instead it's the accumulation of how another civilization nearby refer them to. \emph{Krivittè,} (one) who hides something, is a word referring to the main line of tradition of Creeve people. Here I use Creeve, a rough anglicization of Krivittè to refer both for the language and for those who speak it.


%add map later